---
title: "Final Project Draft - Part 2 Prediction"
author: "Jacob Walsh"
date: "December 6, 2016"
output: word_document
---

```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=65),tidy=TRUE, cache=TRUE)
```
#Part 2 Prediction
```{r eval=FALSE}
charity <- read.csv("C:/Penn State MAS/Stat 897D/charity.csv")
#transform data
t.charity=charity
t.charity$avhv=log(charity$avhv)
t.charity$incm=log(charity$incm)
t.charity$inca=log(charity$inca)
t.charity$plow=charity$plow^(1/3)
t.charity$tgif=log(charity$tgif)
t.charity$lgif=log(charity$lgif)
t.charity$rgif=log(charity$rgif)
t.charity$tdon=log(charity$tdon)
t.charity$tlag=log(charity$tlag)
t.charity$agif=log(charity$agif)


#partition data
data.train <- t.charity[t.charity$part=="train",]
x.train <- data.train[,2:21]
c.train <- data.train[,22] # donr
n.train.c <- length(c.train) # 3984
y.train <- data.train[c.train==1,23] # damt for observations with donr=1
n.train.y <- length(y.train)


data.valid <- t.charity[t.charity$part=="valid",]
x.valid <- data.valid[,2:21]
c.valid <- data.valid[,22] # donr
n.valid.c <- length(c.valid) # 2018
y.valid <- data.valid[c.valid==1,23] # damt for observations with donr=1
n.valid.y <- length(y.valid)


data.test <- t.charity[t.charity$part=="test",]
n.test <- dim(data.test)[1] # 2007
x.test <- data.test[,2:21]

x.train.mean <- apply(x.train, 2, mean)
x.train.sd <- apply(x.train, 2, sd)
x.train.std <- t((t(x.train)-x.train.mean)/x.train.sd) # standardize to have zero mean and unit sd
apply(x.train.std, 2, mean) # check zero mean
apply(x.train.std, 2, sd) # check unit sd
data.train.std.c <- data.frame(x.train.std, donr=c.train) # to classify donr
data.train.std.y <- data.frame(x.train.std[c.train==1,], damt=y.train) # to predict damt when donr=1

x.valid.std <- t((t(x.valid)-x.train.mean)/x.train.sd) # standardize using training mean and sd
data.valid.std.c <- data.frame(x.valid.std, donr=c.valid) # to classify donr
data.valid.std.y <- data.frame(x.valid.std[c.valid==1,], damt=y.valid) # to predict damt when donr=1

x.test.std <- t((t(x.test)-x.train.mean)/x.train.sd) # standardize using training mean and sd
data.test.std <- data.frame(x.test.std)
```

##Model Comparison
###Linear Regression
```{r eval=FALSE}

library(leaps)
model.ls1 <- lm(damt ~ reg1 + reg2 + reg3 + reg4 + home + chld + hinc + genf + wrat + 
                  avhv + incm + inca + plow + npro + tgif + lgif + rgif + tdon + tlag + agif, 
                data.train.std.y)

pred.valid.ls1 <- predict(model.ls1, newdata = data.valid.std.y) # validation predictions
lm.mse=mean((y.valid - pred.valid.ls1)^2) # mean prediction error
sd((y.valid - pred.valid.ls1)^2)/sqrt(n.valid.y) # std error

#best subsets
model.bestsub=regsubsets(damt~., data=data.train.std.y, nvmax=20)
test.mse=model.matrix(damt~.,data=data.valid.std.y)
val.errors=rep(NA,20)
for(i in 1:20){
coefi=coef(model.bestsub,id=i)
pred=test.mse[,names(coefi)]%*%coefi
val.errors[i]=mean((data.valid.std.y$damt-pred)^2)
}
val.errors[which.min(val.errors)]
which.min(val.errors)
coefficients(model.bestsub, which.min(val.errors)) #Coefficients for the "best" model.  There is no reduction in the amount of predictors to minimize the mean prediction error, all 20 predictors are included.

plot(model.bestsub, scale="bic")

model.bestfwd=regsubsets(damt~., data=data.train.std.y, method="forward")
plot(model.bestfwd, scale="bic")

model.bestbwd=regsubsets(damt~., data=data.train.std.y, nvmax=20, method="backward")
plot(model.bestbwd, scale="bic")

#bic for best subsets and backward are the same 10 predictors reg3+reg4+home+chld+hinc+incm+tgif+lgif+rgif+agif
lm8.mse=val.errors[8]
#bic for the best foward regressions is minimized at 8 predictors
lm10.mse=val.errors[10]
#mSE isn't any lower
```

###Ridge and Lasso
```{r eval=FALSE}
library(glmnet)
grid=10^seq(10,-2,length=100)
xtrain=model.matrix(damt~., data=data.train.std.y)
ytrain=data.train.std.y$damt
xtest=model.matrix(damt~., data.valid.std.y)
ytest=data.valid.std.y$damt

model.ridge=glmnet(xtrain,ytrain,alpha=0, lambda=grid)
cv.ridge=cv.glmnet(xtrain,ytrain,alpha=0)
pred.ridge=predict(model.ridge, s=cv.ridge$lambda.min, newx=xtest)
ridge.mse=mean((pred.ridge-ytest)^2)

model.lasso=glmnet(xtrain,ytrain,alpha=1, lambda=grid)
cv.lasso=cv.glmnet(xtrain,ytrain,alpha=1)
pred.lasso=predict(model.lasso, s=cv.lasso$lambda.min, newx=xtest)
lasso.mse=mean((pred.lasso-ytest)^2)

#ridge and lasso aren't performing any better than plain old least squares.
```

###PCR and PLS
```{r eval=FALSE}

#pcr
library(pls)
set.seed(1)
pcr.fit=pcr(damt~.,data=data.train.std.y, scale=TRUE, validation = "CV")
validationplot(pcr.fit,val.type="MSEP")

#not much gained after 5, next slight dip is at 13.
pcr.pred=predict(pcr.fit,data.valid.std.y, ncomp=5)
pcr5.mse=mean((pcr.pred-y.valid)^2)
pcr.pred=predict(pcr.fit,data.valid.std.y, ncomp=13)
pcr13.mse=mean((pcr.pred-y.valid)^2)

pls.fit=plsr(damt~.,data=data.train.std.y, scale=TRUE, validation="CV")
validationplot(pls.fit, val.type="MSEP")

pls.pred=predict(pls.fit,data.valid.std.y, ncomp=2)
pls.mse=mean((pls.pred-y.valid)^2)  #slight decrease in mse, with only 2 components..
```

###SVM

```{r eval=FALSE}
library(e1071)
svmfit1=svm(damt~., data=data.train.std.y, cost =1, gamma=0.01, kernel = "radial")
svmfit2=svm(damt~., data=data.train.std.y, cost =1, gamma=0.01, epsilon=.2, kernel = "radial") 

svmpred1=predict(svmfit1, newdata=data.valid.std.y)
svmpred2=predict(svmfit2, newdata=data.valid.std.y)

svm1.mse=mean((svmpred1-y.valid)^2)
svm2.mse=mean((svmpred2-y.valid)^2) #best so far
```

###Random Forests
```{r eval=FALSE}
library(randomForest)
set.seed(1)
bag.charity=randomForest(damt~.,data=data.train.std.y, mtry=20, importance=TRUE)
pred.valid.bag=predict(bag.charity, newdata=data.valid.std.y)
bag.mse=mean((pred.valid.bag-y.valid)^2)

set.seed(1)
rf.charity=randomForest(damt~.,data=data.train.std.y, mtry=10, ntree=25, importance=TRUE)
pred.valid.rf=predict(rf.charity, newdata=data.valid.std.y)
rf.mse=mean((pred.valid.rf-y.valid)^2)
```

###GBM
I started out by manually changing the parameters using the example in ISLR.  After some research, I found a package that does this for you.  It's called caret and it did better than any other model by far.
```{r eval=FALSE}
library(gbm)
boost.mod=gbm(damt~.,data=data.train.std.y, distribution="gaussian", n.trees=3000, interaction.depth=4)
pred.valid.gbm=predict(boost.mod, newdata=data.valid.std.y, n.trees=1000)
gbm1.mse=mean((pred.valid.gbm-y.valid)^2)

boost.mod=gbm(damt~.,data=data.train.std.y, distribution="gaussian", n.trees=5000, interaction.depth=6)
pred.valid.gbm=predict(boost.mod, newdata=data.valid.std.y, n.trees=5000)
gbm2.mse=mean((pred.valid.gbm-y.valid)^2)
```

```{r eval=FALSE}
library(caret)
set.seed(1)
gbmfit<- train(damt~., data=data.train.std.y, method='gbm')
```
####Optimal GBM
Using the package caret I tuned the parameters to the following
```{r eval=FALSE}
gbmfit
set.seed(1)
boost.mod=gbm(damt~.,data=data.train.std.y, distribution="gaussian", n.trees=150, interaction.depth=3, shrinkage=0.1, n.minobsinnode=10)
pred.valid.gbm=predict(boost.mod, newdata=data.valid.std.y, n.trees=150)
gbmoptimal.mse=mean((pred.valid.gbm-y.valid)^2)
```

##Results
```{r eval=FALSE}
lm.mse
lm8.mse
lm10.mse
ridge.mse
lasso.mse
pcr5.mse
pcr13.mse
pls.mse
svm1.mse
svm2.mse
bag.mse
rf.mse
gbmoptimal.mse
```
